---
title: "Mapping of the Sustainable Food Production Group members"
author: "Ido Bar"
date: "29/07/2021"
output: 
    html_document:
      css: "style/style.css"
      toc: true
      toc_float: true
      toc_depth: 3
      highlight: pygments
      number_sections: false
      code_folding: hide
csl: style/springer-basic-improved-author-date-with-italic-et-al-period.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
pacman::p_load(tidyverse, captioner, ggmap, glue, geosphere,
               pheatmap, kableExtra, DT, scales, formattable, xml2,
               leaflet, leafpm)
# countrycode
# See .Renviron for Google API key
# load function to convert address to coordinates (using Google API)
source("src/getGeo.R")
# DataTable options
options(DT.options = list(pageLength = 15, searching = FALSE))
# Font Format
custom_font="consolas"
fontFmt = function(x,font="consolas"){
  #outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
  #if (outputFormat == 'html')
  formatted_text <- sprintf("<font face='%s'>%s</font>",font,x)
  return(formatted_text)
}

```

```{r captions, include=FALSE}
figs <- captioner(prefix="Figure")
tbls <- captioner(prefix="Table")
tbls(name="ratios","Aspect ratios of flags of selected countries")
tbls("emoji_flags", "Country name/ISO3166 code to Unicode conversion")
tbls(name="distance", "Distance matrix between group members")
tbls(name="dist_sum", "Summary of closest and furthest distances from each group member original location")
figs(name="heatmap", "Heatmap showing the distance between group members and their corresponding clustering")
```

# Geographical mapping of group members

In our last group lunch we realised that each member of our research group came from a different country, so we wondered who's closest and who's furthest from each other.  This post describes a little toy example of collecting addresses (in this case our research group members) and plotting them on a map, while demonstrating the distances between members.

## Collect addresses

Members of the Sustainable Food Production Group (at Griffith University) provided their hometown and country information, which was manually saved into an excel spreadsheet (in the future please use the following [online form](https://forms.office.com/Pages/ResponsePage.aspx?id=q8h8Wtykm0-_YGZxQEmtYveRT5yQMCdChpx4HeAKmENURDlIOUNWUEIwWVBJVUdPVldHWUZNMEdQSS4u)).  
The address of each member was converted to coordinates using the `geocode()` function from the [ggmap](https://github.com/dkahle/ggmap) package (utilising Google Maps API and parsed based on the script in [Shane Lynn's blog post](https://www.shanelynn.ie/massive-geocoding-with-r-and-google-maps/)). Country names were translated to ISO 3166 2-letter codes (short names) using the same output from the Google API. 


```{r addresses}
# load required packages
pacman::p_load(tidyverse, ggmap, glue, geosphere,
               pheatmap, scales, formattable, xml2,
               leaflet, leafpm, paletteer)
# read in addresses from an excel
addresses <- readxl::read_excel("data/lab_members_addresses.xlsx") %>% 
           mutate(first_name=sub(" .+", "", Name))
update_addresses <- TRUE
if (isTRUE(update_addresses)){
  # Get geodetails (up to 2,500 queries a day), otherwise use getGeo
  geodetails <- geocode(addresses$Address, output = "all", messaging=TRUE)
  names(geodetails) <- addresses$Address
  geodetails %>% discard(~ .$status=="OK") %>% iwalk(~message(glue("Could not find details for {.y}")))
  # geodetails %>% keep(~ "university" %in% flatten_chr(.$results[[1]]$types)) %>% 
  #   walk(~message(glue("{.} is a university")))
  
  results <- geodetails %>% keep(~ .$status=="OK") %>% map_df(~tibble(
    formatted_address=.$results[[1]]$formatted_address,
    lat=.$results[[1]]$geometry$location$lat,
    long=.$results[[1]]$geometry$location$lng,
    Country=.$results[[1]]$address_components %>% 
      keep(~ "country" %in% flatten_chr(.$types)) %>% flatten() %>% .$long_name,
    country_code=.$results[[1]]$address_components %>% 
      keep(~ "country" %in% flatten_chr(.$types)) %>% flatten() %>% .$short_name,
    status=.$status)) %>% mutate(query=names(geodetails %>% keep(~ .$status=="OK")))
  geo_df <- results %>% 
  # geo_df <- addresses$Address %>% map_dfr(~getGeoDetails(.x)) %>% 
    left_join(addresses, ., by=c("Address"="query")) %>% 
    write_csv("data/lab_members_geodata.csv")
} else geo_df <- read_csv("data/lab_members_geodata.csv")
```

## Download flags

### Rectangular flags
Flags of the representative countries were downloaded to be used as map markers, both in rectangular and in a round emoji shape.  The flags were downloaded from the [Google region flags](https://github.com/google/region-flags) GitHub repository and stored locally in `svg` format.  

```{r rect_flags}
# download flags
update_flags <- TRUE
walk(geo_df$country_code, ~{
  if (!file.exists(glue("img/flags/{.x}.svg")) | isTRUE(update_flags)) {
    download.file(glue('https://raw.githubusercontent.com/google/region-flags/gh-pages/svg/{.x}.svg'), 
                       glue("img/flags/{.x}.svg"))
  } else {
    message(glue("File img/flags/{.x}.svg exists"))
    }
})
```

Note that I used the `walk()` function from the [purrr](https://purrr.tidyverse.org/) package to iterate over the filenames and download them, without returning any value. `walk()` is similar to the `map()` functions from the same package, which in turn provide functional programming alternatives to the `.apply()` functions in base R (see useful [purrr tutorial](https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html) and some [extra tips](https://www.hvitfeldt.me/blog/purrr-tips-and-tricks/#purrr-tips-and-tricks)) 

Special attention was needed when using the rectangular flags as markers, since each flag has a different height:width aspect ratio and a fixed width and height will distort some. The solution was to parse the `svg` files, which are basically in `xml` format, using the [xml2](https://xml2.r-lib.org/) package and calculate the aspect ratio from their width and height (see the function below and results in `r tbls(name="ratios", display="cite")`).

```{r aspect_ratio}
getAspectRatioSVG <- function(img_file){
  require(xml2)
  # img_file=rect_flags$flag_file[3]
  img_info <- read_xml(img_file) 
  aspect_ratio <- as.numeric(xml_attr(img_info, "width"))/as.numeric(xml_attr(img_info, "height"))
  if (length(aspect_ratio)>0 & is.numeric(aspect_ratio) & aspect_ratio>0) return(aspect_ratio)
  message(sprintf("File %s is not a valid SVG file or does not contain dimensions", img_file))
}
# specify a base marker height
base_height <- 15
# get flags aspect ratios from SVG files and calculate appropriate width
rect_flags <-  geo_df %>% select(country_code) %>%
  mutate(flag_file=glue("./img/flags/{country_code}.svg"),
         Ratio=map_dbl(flag_file, ~getAspectRatioSVG(.x)),
         flag_height=base_height, flag_width=base_height*Ratio)
```
      

```{r aspect_ratio_table, echo=FALSE}
# datatable(as.data.frame(rect_flags), caption=tbls("ratios"), rownames = FALSE)
rect_flags %>% mutate_if(is.double, ~round(.x, 3)) %>% 
  kable(caption = tbls("ratios")) %>% 
  kable_styling(c("striped", "condensed"), full_width = F)

```

>An alternative is to parse an online table containing the [flags aspect ratios](https://en.wikipedia.org/wiki/List_of_aspect_ratios_of_national_flags) from wikipedia (using [htmltab](https://github.com/crubba/htmltab) package) and use the ratios to calculate the appropriate width for each flag, but it was more cumbersome and difficult to match countries just by name.

Once I have the dataframe with the flags information, I could iterate through it (this time using `pmap()` function from [purrr](https://purrr.tidyverse.org/)) to create an icon-set that I could use as markers in the map.


I then plotted the map with [leaflet](https://rstudio.github.io/leaflet/) package, using the `Esri.NatGeoWorldMap` tiles as a basemap, along with optional scale bar and distance/area measurement tool (the map is interactive, please play with it and hover over the markers).


```{r flags_map}
# Create a named list of icons/flags
flagIcons <-  rect_flags[c("flag_file", "flag_width", "flag_height")] %>%
  pmap(~makeIcon(..1, ..1,  iconWidth = ..2, iconHeight = ..3,
                 iconAnchorX = 0, iconAnchorY = 0)) %>%
  setNames(., rect_flags$country_code)
# make it an iconList
attr(flagIcons, "class") <- "leaflet_icon_set"

# Create the leaflet map
leaflet(geo_df) %>% addProviderTiles("Esri.NatGeoWorldMap") %>%
  addMarkers(popup = ~Name, label = ~Name, 
             icon = ~flagIcons[country_code]) %>% # ~flagIcons[country_code]
  addScaleBar(position = "bottomright",
              options = scaleBarOptions(imperial = FALSE)) %>% 
  addMeasure(primaryLengthUnit = "kilometers")
```

### Emoji flags
An alternative visualisation was then used, trying to mimick the [ggflags](https://github.com/rensa/ggflags) behaviour and use round flag emojis from the discontinued [EmojiOne Colour Font](https://github.com/eosrei/emojione-color-font) repository.  In this case, I faced another challenge, trying to download just the `svg` files of the required flags out of the thousands of emojis that use their Unicodes as file names. This time, [Tim Whitlock's blog](https://apps.timwhitlock.info) came to my rescue with an organised [Country-ISO3166-Unicode table](https://apps.timwhitlock.info/emoji/tables/iso3166) that I could parse with [htmltab](https://github.com/crubba/htmltab) and construct the filenames to be downloaded (see `r tbls(name="emoji_flags", display="cite")`).    

```{r emoji_flags}
# Use round emoji flags as markers
round_flags <- htmltab::htmltab("https://apps.timwhitlock.info/emoji/tables/iso3166") %>% as_tibble() %>% 
  mutate(flag_file=tolower(sub(" ", "-", 
                              gsub("U+", "", Unicode, fixed = TRUE)))) %>%
  select(Country=Name, country_code=ISO, flag_file) %>% 
  inner_join(geo_df %>% select(country_code), .) 

walk2(unique(round_flags$flag_file), unique(round_flags$country_code), ~{
  if (!file.exists(glue("img/emojis/{.y}.svg")) | isTRUE(update_flags)) {
    download.file(glue('https://raw.githubusercontent.com/eosrei/emojione-color-font/master/assets/emojione-svg/{.x}.svg'), glue("img/emojis/{.y}.svg"))
  } else {
    message(glue("File img/emojis/{.y}.svg exists"))
  }
      })
```

```{r emoji_table, echo=FALSE}
# datatable(as.data.frame(round_flags), caption=tbls("emoji_flags"), rownames = FALSE)
round_flags %>% select(Country, country_code, flag_file) %>% 
  kable(caption=tbls("emoji_flags")) %>% 
  kable_styling(c("striped", "condensed"), full_width = F) %>% 
  column_spec(c(1,3), width = "15em") 

```

I plotted the same map as before, but using the emoji flags as markers and a different basemap (`Esri.WorldStreetMap`) for the interest (see complete list of available maps [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html) and note that some providers require registration or setting up an API key, see [details](https://github.com/leaflet-extras/leaflet-providers))

```{r emoji_map}
base_height <- 15
# Create a named list of icons/flags
emojiIcons <- icons(glue("img/emojis/{round_flags$country_code}.svg"),
                    iconHeight = base_height, iconWidth = base_height,
                    iconAnchorX = 0, iconAnchorY = 0)

# Create the leaflet map
leaflet(geo_df) %>% addProviderTiles("Esri.WorldStreetMap") %>% # 
  addMarkers(popup = ~Name, label = ~Name, 
             icon = emojiIcons) %>% 
  addScaleBar(position = "bottomright",
              options = scaleBarOptions(imperial = FALSE)) %>% 
  addMeasure(primaryLengthUnit = "kilometers")
```

## Calculate distances

Once the map were plotted, I wanted to get a better and more accurate understanding of the distances between the hometowns of the group members. That was easily calculated using `distm()` function (with the `distVincentyEllipsoid` method) from the [geosphere](https://cran.r-project.org/web/packages/geosphere/index.html) package.

The distanced were then plotted using [pheatmap](https://github.com/raivokolde/pheatmap) package, that offers lots of customisation and annotation options (see [Dave Tang's Blog](https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/) for usage examples), as well as plotting a dendrogram that shows the relatedness between members (`r figs(name="heatmap", display="cite")`).

`r figs("heatmap")`
```{r dist_heatmap}
# create distance matrix
dist_mat <- distm(geo_df[c('long','lat')], geo_df[c('long','lat')], 
             fun=distVincentyEllipsoid)/1000
dimnames(dist_mat) <- list(geo_df$first_name, geo_df$first_name)
# set the diagonal to NA
dist_mat[dist_mat==0] <- NA

pheatmap(dist_mat, cutree_rows = 3, 
         color = paletteer_c("viridis::inferno", 30, direction = -1),
         cutree_cols = 3)

```

The distances can actually be observed in the matrix:

```{r dist_matrix, echo=FALSE}
dist_mat %>% as_tibble() %>% 
  mutate_all(~round(., 2)) %>% 
  # mutate_all(~color_bar("lightgreen")(.)) %>% 
  mutate_all(~cell_spec(., color = "white", bold = T,
    background = spec_color(., end = 0.9, option = "A",
                            direction = -1))) %>%
  mutate(rownames=colnames(.)) %>% column_to_rownames("rownames") %>% 
  kable(escape = FALSE, caption = tbls("distance")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)
```

Finally, we can look at the closest colleague of each member to decide who's hometown is the most isolated (`r tbls(name="dist_sum", display="cite")`).

```{r min_dist}
# change into a long format
long_dist <- as.data.frame(dist_mat) %>% rownames_to_column("Name") %>% 
  gather("Colleague", "Distance", 2:ncol(.)) 
# find closest location for each member
closest <- long_dist %>%  filter(!is.na(Distance)) %>% group_by(Name) %>% slice_min(Distance) %>% 
  mutate(Closest=glue("{Colleague} ({round(Distance, 0)} km)")) %>% select(Name, Closest)
   # `Distance to closest colleague`
# find furthest location from each member and combine tables
summary_table <- long_dist %>%  filter(!is.na(Distance)) %>% group_by(Name) %>% 
  slice_max(Distance) %>% 
  mutate(Furthest=glue("{Colleague} ({round(Distance, 0)} km)"))  %>%  # `Distance to furthest colleague`
  select(Name, Furthest) %>% inner_join(closest, .) %>% 
  setNames(., c("Name", "Closest member", "Furthest member"))
```
```{r dist_sum_table, echo=FALSE}
summary_table %>% 
  # mutate_if(is.numeric, ~round(., 2)) %>% 
  # # mutate_all(~color_bar("lightgreen")(.)) %>% 
  # mutate_if(is.numeric, ~cell_spec(., color = "white", bold = T,
  #   background = spec_color(., end = 0.9, option = "A",
  #                           direction = -1))) %>%
  kable(escape = FALSE, caption = tbls(name="dist_sum")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)
```

# Alternatives

A few other mapping packages are available as an alternative to [leaflet](https://rstudio.github.io/leaflet/), such as [mapview](https://r-spatial.github.io/mapview/) (along with its extension package to leaflet [leafem](https://github.com/r-spatial/leafem)).   [googleway](http://symbolixau.github.io/googleway/) is another package that allows usage of Google Maps and the entire Google APIs (places, directions, roads, distances, geocoding, elevation and timezone). 