---
title: "Presentation Schedule"
author: "Ido Bar"
date: "15/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, lubridate, gt, janitor)
```

## Presentation Schedule

Each meeting, a member of the team presents a quick presentation on some scientific topic, which can be their work, other published research/methods or anything else that's relevant for our research (can be about academic writing, computer skills, etc.).  
Another member of the team will chair the presentation, i.e, he/she will introduce the speaker, manage questions and engage other members in the conversation.

To determine the order of presentations and chairperson, we will use the following approach:

```{r random-speaker}
start_date <- dmy("23/06/2022")
members <- c("Fawad", "Matin", "Gurpreet", "Shane", "Mel", "Moutoshi",  "Josh", "Hayley",
             "Henrietta", "Sam", "Rebecca", "Mahmuda", "Ido", "Chat")
set.seed(15)
presentation_schedule <- tibble(presenter = sample(members), chairperson = sample(members), 
                                date = seq(start_date, by = "2 week",
                                           length.out = length(members)))
presentation_schedule %>% 
  clean_names("title") %>% 
  gt() %>% 
  fmt_date(Date, date_style = "day_month_year")

```
